---
title: "What happened in TX-22?"
author: Joshua Kravitz
date: '2021-01-17'
slug: []
categories: []
tags: []
subtitle: ''
summary: ''
authors: []
lastmod: '2021-01-17T21:05:05-08:00'
featured: no
draft: no
image: 
  caption: ''
  focal_point: ''
  preview_only: yes
projects: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error = FALSE)
```

This is a test post!



```{r}
library(tidyverse)
library(ggtext)

color_dem <- "#0015BC"
color_gop <- "#E9141D"
```

```{r}
pct_results <- 
  read_csv(
    "data/1103_results_by_pct.csv",
    col_types = cols(
      county = col_character(),
      office = col_character(),
      .default = col_number()
    )
  ) 

pct_results <- 
  pct_results %>% 
  semi_join(
    pct_results %>% 
      filter(office == "tx22") %>% 
      select(county, pct),
    by = c("county", "pct")
  ) %>% 
  mutate(votes = dem + gop + other) %>% 
  mutate(
    dem_vote_share = dem / votes,
    gop_vote_share = gop / votes,
    mov = dem_vote_share - gop_vote_share
  ) %>% 
  relocate(mov, dem_vote_share, gop_vote_share, .after = pct)

pct_demographics <- read_csv(
  "data/1103_demographics_by_pct.csv",
  col_types = cols(
    county = col_character(),
    .default = col_double()
  )
)

pct_scores <- read_csv(
  "data/precinct_scores.csv",
  col_types = cols(
    state_code = col_character(),
    .default = col_double()
  )
)
```

```{r}
  # It seems results for these precincts just don't exist from 2018.
  # I'm not sure why, but I checked:
  #   1) `democrats.election_results.g2018_tsmart_precinct` in BQ
  #   2) MIT Election Lab data: https://github.com/MEDSL/2018-elections-official/blob/master/precinct_2018.zip
  # They both didn't have these precincts
  # filter(!is.na(mov_2020), !is.na(mov_2018))

pct_tx22 <- 
  pct_results %>% 
    filter(office == "tx22")
  
pct_tx22_wide <- 
  pct_tx22 %>% 
    pivot_wider(
      id_cols = c(county, pct),
      names_from = year,
      values_from = c(votes, mov)
    ) %>% 
    left_join(pct_demographics, by = c("county", "pct")) %>% 
    mutate(
      turnout_2020 = votes_2020 / registered_voters_2020,
      turnout_2018 = votes_2018 / registered_voters_2018,
      turnout_2020 = if_else(turnout_2020 > 1, NA_real_, turnout_2020),
      turnout_2018 = if_else(turnout_2018 > 1, NA_real_, turnout_2018)
    ) 
```

```{r}
pct_results %>% 
  group_by(year, office) %>% 
  summarize(
    votes = sum(votes), 
    dem_margin = round((sum(dem) / sum(votes) - sum(gop) / sum(votes)) * 100, 2)
  ) %>% 
  knitr::kable()
```

## How did we fare compared to 2018?

```{r}
pct_tx22_wide %>% 
  ggplot(aes(mov_2020 - mov_2018)) +
  geom_histogram(binwidth = 0.025) +
  geom_vline(xintercept = 0, color = "red", alpha = 0.5, size = 2) +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1)) +
  theme_light() +
  labs(
    title = "TX-22: '18 to '20 Margin of Victories by PCT",
    x = "2018-2020 change in Democratic vote margin"
  )
```

We lost support in many precincts: what happened?

To investigate, we'll perform an Ecological Study at the precinct level, for each of 4  ethnicities modeled in the voter file: White, Black, Asian, and Hispanic. 

```{r}
margin_change_by_demographic <- function(...) {
  pct_plot(
    ..., 
    y_var = mov_2020 - mov_2018,
    y_lab = "2018-2020 change in\nDemocratic vote margin, %"
  )
}

pct_plot <- function(
  df, x_var, y_var,
  title, x_lab, y_lab,
  x_min = 0, x_max = 1, x_jump = 0.05, 
  y_min = -1, y_max = 1, y_jump = 0.05,
  href = TRUE
) {
  x_var <- enquo(x_var)
  y_var <- enquo(y_var)
  
  g <- 
    df %>% 
    ggplot(aes(!!x_var, !!y_var))
  
  if (href) {
    g <- g +  geom_hline(yintercept = 0, color = color_gop, alpha = 0.2, size = 2)
  }
  
  g +
    geom_point(
      aes(size = votes_2020, fill = mov_2020),
      # alpha = 0.5,
      shape = 21
    ) +
    geom_smooth() +
    scale_x_continuous(
      breaks = seq(x_min, x_max, x_jump),
      minor_breaks = FALSE,
      labels = scales::label_percent(accuracy = 1, suffix = "")
    ) +
    scale_y_continuous(
      breaks = seq(y_min, y_max, y_jump), 
      minor_breaks = FALSE,
      labels = scales::label_percent(accuracy = 1, suffix = ""),
    ) +
    scale_fill_gradient2(
      low = color_gop, 
      high = color_dem,
      labels = scales::label_percent(accuracy = 1, suffix = "")
    ) +
    scale_size(range = c(1, 10)) +
    labs(
      title = title,
      x = x_lab,
      y = y_lab,
      size = "*2020 votes*",
      fill = "*2020 Democratic margin, %*"
    ) +
     guides(
      fill = guide_colorbar(
        title.position = "top",
        barwidth = 8,
        barheight = 0.25,
        raster = FALSE,
        nbin = 8,
        ticks = FALSE,
      ),
      size = guide_legend(title.position = "top")
    ) +
    theme_minimal() + 
    theme(
      legend.position = "bottom",
      legend.title = element_markdown(size = 8),
      legend.text = element_text(size = 6)
    )
}
```

## 2018-2020 change in Democratic margin, by ethnicity

```{r}
pct_tx22_wide %>% 
  margin_change_by_demographic(
    model_white / registered_voters_2020, 
    "White", 
    "White voters, %",
    x_jump = 0.1
  )
```


```{r}
pct_tx22_wide %>% 
  margin_change_by_demographic(
    model_aapi / registered_voters_2020, 
    "AAPI", 
    "AAPI voters, %",
    x_jump = 0.1
  )

pct_tx22_wide %>% 
  margin_change_by_demographic(
    model_black / registered_voters_2020, 
    "Black", 
    "Black voters, %"
  )

pct_tx22_wide %>% 
  margin_change_by_demographic(
    model_latinx / registered_voters_2020, 
    "Latinx", 
    "Latinx voters, %"
  )
```
Some key takeaways:

1. We lost support in precincts with a higher proportion of AAPI voters.
2. We lost support in precincts with a lower proportion of white voters.
3. In precincts with a higher proportion of white voters, support remained steady compared to 2018, though we still suffered big losses in 2020.
4. Precincts with a higher proportion of black voters supported us.
5. Support among high-density latinx precincts was generally worse this cycle, though overall these precincts were a mixed bag (some went for Nehls, some went for Kulkarni).

AAPI is most interesting, given it was a major focus of our campaign. Let's dive in.


## What happened to the AAPI population?

# ```{r}
# pct_tx22_wide %>% 
#   margin_change_by_demographic(
#     desi / registered_voters_2020, 
#     "High-density Desi precincts lost support", 
#     "Desi voters, %"
#   )
# ```

```{r}
pct_tx22_wide %>% 
  mutate(
    pct_desi = desi / registered_voters_2020,
    highlight = pct_desi >= 0.1
  ) %>% 
  ggplot(aes(mov_2018, mov_2020 - mov_2018)) +
  geom_hline(yintercept = 0, color = "gray", alpha = 0.5, size = 1) +
  geom_vline(xintercept = 0, color = "gray", alpha = 0.5, size = 1) +
  geom_point(
    aes(
      size = votes_2020, 
      color = pct_desi, 
      alpha = highlight
    ), 
    shape = 20
  ) +
  geom_text(aes(
    label = if_else(highlight, as.character(pct), "")), 
    size = 1.5
  ) +
  annotate(
    "text", 
    x = Inf, 
    y = 0, 
    label = c("Better 2020 margin", "Worse 2020 margin"),
    size = 2,
    color = c(color_dem, color_gop),
    hjust = 1.05,
    vjust = c(-1, 1.8)
  ) +
  scale_alpha_discrete(range = c(0.1, 0.5)) +
  scale_size(breaks = seq(0, 10000, 2500), range = c(1, 15)) +
  scale_color_viridis_c(label = scales::percent_format(accuracy = 1)) +
  scale_x_continuous(
    breaks = seq(-1, 1, 0.2), 
    minor_breaks = FALSE,
    labels = scales::percent_format()
  ) +
  scale_y_continuous(
    breaks = seq(-1, 1, 0.05), 
    minor_breaks = FALSE,
    labels = scales::percent_format(accuracy = 1)
  ) +
  guides(
    alpha = FALSE, 
    size = FALSE, 
    color = guide_colorbar(
      barwidth = 0.25,
      barheight = 9,
      raster = FALSE,
      nbin = 8,
      ticks = FALSE
    )
  ) +
  labs(
    title = "Change in precinct margin (precincts with > 10% desis are highlighted)",
    subtitle = "High-density Desi precincts lost support",
    x = "2018 Democratic vote margin",
    y = "2018-2020 change in Democratic vote margin",
    color = "% Desi",
    caption = "Source: DNC ('18 data), FB/Brazoria/Harris Board of Elections ('20 data)"
  ) +
  theme_minimal() + 
  theme(plot.caption = element_text(size = 5))
```
Both these plots basically show the same thing -- high-density Desi precincts lost support.
Let's see if Hindus and Muslims behaved any differently.

```{r}
pct_tx22_wide %>% 
  mutate(
    better = mov_2020 - mov_2018 > 0,
    frac_muslim = muslim / registered_voters_2020,
    frac_hindu = hindu / registered_voters_2020
  ) %>%
  ggplot(aes(frac_muslim, frac_hindu)) +
  geom_point(
    aes(
      size = abs(mov_2020 - mov_2018), fill = better),
    alpha = 0.3,
    shape = 21
  ) +
  scale_x_continuous(
    breaks = seq(0, 1, 0.05), 
    minor_breaks = FALSE,
    labels = scales::percent
  ) +
  scale_y_continuous(
    breaks = seq(0, 1, 0.05), 
    minor_breaks = FALSE,
    labels = scales::percent
  ) +
  guides(
    fill = FALSE
  ) +
  labs(
    title = "Change in precinct margin: Hindu and Muslim populations",
    x = "% Muslim",
    y = "% Hindu",
    # size = "No. 2020 Voters",
    size = "Change in Margin",
    caption = "Source: DNC ('18 data), FB/Brazoria/Harris Board of Elections ('20 data)"
  ) +
  theme_minimal() +
  theme(plot.caption = element_text(size = 5), legend.position = "bottom")
```
They're pretty well-correlated, so not much to find here.

So it looks like we did better with whites than minorities (not in absolute terms, just compared to 2018).
To Note: This assumes static demographics for the precincts since 2018 -- so e.g., if more minorities entered these
white-heavy communities (even if they were still white-heavy), it's possible the margins went up not because of change
in voting behavior of white voters, but rather because of a change in demographics.

Same thing for non-white communities: maybe votes decreased for Sri because of a population shift away from those communities
of color ... that feels unlikely but possible.

If I can get 2018 demographic data, can I do a difference in difference?

Sri says generally non-white proportion of electorate increased ... so increase in white precinct support may just come
from more non-white voters in those white-heavy precincts (compared to 2018).


## How did Kulkarni compare to Biden and Hegar?


```{r}
# Maybe something? Shows gap in dem margin b/t president and senate/congress that's
# largest in high-density AAPI precincts.
pct_results %>% 
  filter(year == 2020) %>% 
  left_join(pct_demographics, by = c("county", "pct")) %>% 
  ggplot(aes(model_aapi / registered_voters_2020, mov)) +
  geom_smooth(aes(group = office, color = office), alpha = 0.2) +
  geom_point(aes(fill = office), alpha = 0.5, shape = 21) +
  labs(
    title = "Margin of Victory "
  ) +
  theme_minimal()
```

```{r}
pct_results %>% 
  filter(year == 2020) %>% 
  group_by(year, office) %>% 
  summarize(
    gop = sum(gop), 
    dem = sum(dem),
    other = sum(other),
    .groups = "drop"
  ) %>% 
  transmute(
    office,
    votes = dem + gop + other,
    dem_vote_share = dem / votes * 100,
    gop_vote_share = gop / votes * 100,
    other_vote_share = other / votes * 100,
    mov = (dem_vote_share - gop_vote_share)
  )
```


```{r}
df_annotation <- 
  tibble(
    county = factor("Fort Bend", levels = c("Fort Bend", "Brazoria", "Harris")),
    year = "2020 Kulkarni v. Hegar"
  )
pct_results %>% 
  # filter(year == 2018 | office == "president") %>% 
  pivot_wider(
    id_cols = c(county, pct),
    names_from = c(office, year),
    names_prefix = "mov_",
    values_from = mov
  ) %>% 
  left_join(pct_demographics, by = c("county", "pct")) %>% 
  mutate(
    mov_dropoff_2020 = mov_tx22_2020 - mov_senate_2020,
    mov_dropoff_2018 = mov_tx22_2018 - mov_senate_2018
  ) %>% 
  filter(!is.na(mov_dropoff_2020), !is.na(mov_dropoff_2018)) %>% 
  pivot_longer(
    cols = starts_with("mov_dropoff"),
    names_to = "year",
    names_prefix = "mov_dropoff_",
    values_to = "mov_dropoff"
  ) %>% 
  mutate(
    year = fct_recode(
      year, 
      "2018 Kulkarni v. O'Rourke" = "2018",
      "2020 Kulkarni v. Hegar" = "2020"
    ),
    county = fct_relevel(county, "Fort Bend", "Brazoria", "Harris")
  ) %>% 
  ggplot() +
  geom_histogram(aes(mov_dropoff), binwidth = 0.01) +
  geom_vline(
    aes(xintercept = avg, color = avg > 0),
    alpha = 0.5,
    size = 1,
    data = function(df) {
      df %>%
        group_by(county, year) %>% 
        summarize(avg = median(mov_dropoff), .groups = "drop")
    }
  ) +
  geom_vline(xintercept = 0, color = "grey", alpha = 0.5, size = 1) +
  geom_text(
    x = -6, y = 18,
    label = "Median ΔMoV\nacross precincts",
    size = 2.5,
    data = df_annotation,
  ) +
  geom_segment(
    x = -3.5, y = 18.5, xend = -2, yend = 15.5,
    arrow = arrow(length = unit(0.2, "cm")),
    data = df_annotation
  ) +
  scale_x_continuous(
    breaks = seq(-1, 1, 0.02),
    minor_breaks = FALSE,
    labels = scales::label_percent(accuracy = 1)
  ) +
  scale_color_manual(values = c(color_gop, color_dem)) +
  facet_grid(rows = vars(county), cols = vars(year)) +
  labs(
    title = str_glue(
      "Did the median TX-22 precinct perform **<span style='color:{color_dem}'>better</span>** or **<span style='color:{color_gop}'>worse</span>** for Kulkarni <br/>than for the up-ballot Senate race?"
    ),
    subtitle = "In 2018, median **Democratic margin** is close across all three counties.<br/>In 2020, median **Margin of Victory (MoV)** behind in Fort Bend: Nehls' name recognition, or something else?",
    x = "Difference in Democratic margin (Kulkarni minus Senate)",
    caption = "Source: DNC ('18 data), FB/Brazoria/Harris Board of Elections ('20 data)"
  ) +
  guides(color = FALSE) +
  theme_bw() +
  theme(
    plot.title = element_markdown(),
    plot.subtitle = element_markdown(size = 8),
    plot.caption = element_text(size = 5),
  )
```

```{r}
pct_results %>% 
  left_join(pct_demographics, by = c("county", "pct")) %>% 
  filter(year == "2020", office == "tx22") %>% 
  group_by(year, office, county) %>% 
  summarize(sum(desi, na.rm = TRUE)/sum(registered_voters_2020, na.rm = TRUE))
```
```{r}
pct_results %>% 
  pivot_wider(
    id_cols = c(county, pct),
    names_from = c(office, year),
    names_prefix = "votes_",
    values_from = votes
  ) %>% 
  mutate(undervotes = votes_president_2020 - votes_tx22_2020) %>% 
  left_join(pct_tx22_wide, by = c("county", "pct")) %>% 
  pct_plot(
    model_white / registered_voters_2020, 
    undervotes / votes_president_2020,
    title = "Ballot drop-off more common in precincts with more non-white voters",
    x_lab = "White Voters, %",
    y_lab = "Undervotes, %",
    y_jump = 0.01,
    href = FALSE
  ) +
  labs(
    subtitle = "Undervotes: fraction of presidential voters who didn't vote for their U.S. House representative"
  )
  # theme(title = elemnt_text(size = 5))
```

```{r}
pct_results %>% 
  pivot_wider(
    id_cols = c(county, pct),
    names_from = c(office, year),
    names_prefix = "votes_",
    values_from = votes
  ) %>% 
  summarize(sum(votes_president_2020))
  
```


---

Theories
---------
Did Hindus & Muslims vote for Nehls, Leblanc, or not vote at all?
Muslims might have voted for Leblanc (didn't like Troy). Hindus likely voted for Nehls.

phone banks -- who we dialed, who answered the phone (demographics),

how many voted from each group, what was their propensity dem or gop, % of that demographic in TX-22
